When does vapor pressure deficit drive or reduce evapotranspiration?

### 2017-12-14 doc/agu_2017/vpd.tex
before first run through
small edits

\usepackage[english]{babel}
\usepackage[export]{adjustbox}
% \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{relsize}
\usepackage{textcomp}
\usepackage[absolute,overlay]{textpos}
\graphicspath{{figures/}}
\DeclareMathOperator{\sign}{sign}

\usepackage{caption}
\captionsetup{font=scriptsize, labelfont=scriptsize}

@@

\title[Your Short Title]{When does vapor pressure deficit drive or reduce evapotranspiration?}
\author{Adam Massmann,  Pierre Gentine and Changjie Lin}
\institute{AGU Fall Meeting}
\date{December 14th, 2017}

@@

\section{Method}
\begin{frame}{Develop a theory to quantify VPD effects}
  \begin{itemize}
  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system.
    \begin{itemize}
    \item While simplifying the system aids intrinsic understanding, simplifying assumptions sacrifice physical realism.
    \end{itemize}
  \item Because we run the risk of over-simplification, we will use FLUXNET2015 data to test how well our theory matches the data. 
  \end{itemize}
\end{frame}

@@

\begin{frame}{Use physically reasonable assumptions remove implicit dependence}
  Apply a constant uWUE assumption (conserved within plant type; see \cite{Zhou_2016}):
  \[uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET},\]
  To derive a new form of Penman-Monteith without implicit ET dependence:
  \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)}{ \Delta + \gamma}\]
\end{frame}

@@

    \begin{itemize}
    \item $c_p$ and $R*$ are constants
    \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
    \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type (e.g. grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
    \end{itemize}
    \onslide<4>
    \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]

    \begin{itemize}
    \item $c_p$ and $R*$ are constants
    \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
    \item uWUE and $g1$ are constants within plant type (e.g. grass, crops, deciduous broadleaf forest, evergreen needleleaf forest, shrub)
    \end{itemize}
    So \textbf{within each plant type}, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is approximately \textbf{just a function of VPD}!
  \end{overprint}    
\end{frame}

@@

\begin{frame}{Summary - When does VPD drive or reduce ET?}
  \begin{itemize}
  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmospheric demand dominates).
  \item For forest sites, environmental VPD approximately straddles the critical VPD.
  \item For shrubs environmental VPD never exceeds the critical VPD.
  \item Theory tested poorly with FLUXNET data for for crops and grass. 
  \item All plant types exhibited a response far below that of PET.
  \item The new uWUE-version of Penman-Monteith we derived could be used as a replacement for PET in drought indices over vegetated surfaces.
  \end{itemize}
\end{frame}

>>>>

\usepackage[english]{babel}
\usepackage[export]{adjustbox}
% \usepackage[utf8x]{inputenc}
\usepackage[style=authoryear]{biblatex}
\addbibresource{references.bib}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage{relsize}
\usepackage{textcomp}
\usepackage[absolute,overlay]{textpos}
\graphicspath{{figures/}}
\DeclareMathOperator{\sign}{sign}
\usepackage{caption}
\captionsetup{font=scriptsize, labelfont=scriptsize}

@@

\title[Your Short Title]{When does vapor pressure deficit drive or reduce evapotranspiration?}
\author{Adam Massmann,  Pierre Gentine and Changjie Lin}
\institute{AGU Fall Meeting}
\date{December 14th, 2017}
\titlegraphic{\vspace*{1.0cm}\hspace*{-0.5cm}\includegraphics[width=6.0cm]{cu_logo.pdf}\hspace*{6.00cm}~%
   \includegraphics[width=1.5cm]{nsf.png}
}

@@

\section{Method}
\begin{frame}{Develop a theory to quantify VPD effects}
  \begin{itemize}
  \item The goal is to simply the problem (increase transparency) while still capturing the leading order behavior the system.
    \begin{itemize}
    \item Simplifying the system aids intrinsic understanding, but at a physical realism.
    \end{itemize}
  \item Because we run the risk of over-simplification, we will use FLUXNET2015 data to test how well our theory matches the data.
    \begin{itemize}
    \item We will test in the growing season of 5 plant types: Deciduous broadleaf forest, evergreen needleleaf forest, shrub, grass, and crops.
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Use physically reasonable assumptions remove implicit dependence}
  Apply a constant uWUE assumption (conserved within plant type; see \cite{Zhou_2016}):
  \[uWUE = \frac{GPP \cdot \sqrt{VPD}}{ET},\]
  \begin{overprint}
    \onslide<2>
    To derive a new form of Penman-Monteith without implicit ET dependence:
    \[  ET = \frac{\Delta R + \frac{g_a\; P}{T} \left( \frac{ c_p \, VPD}{R_{air}} -  \frac{\gamma c_s \sqrt{VPD} }{ R* \; 1.6 \text{ uWUE } (1 + \frac{g_1}{\sqrt{VPD}})} \right)}{ \Delta + \gamma}\]
    \end{overprint}
\end{frame}

@@

    \begin{itemize}
    \item $c_p$ and $R*$ are constants
    \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
    \item \textcolor{magenta}{\textbf{uWUE}} and $\textcolor{magenta}{\boldsymbol{g1}}$ are constants within plant type.
    \end{itemize}
    \onslide<4>
    \[\frac{\partial \, ET}{\partial \, VPD} = \text{scaling} \cdot \left(\frac{ c_p}{R_{air}} - \frac{\gamma c_s }{1.6 \; R*\; \text{ uWUE }} \left( \frac{2 g_1 + \sqrt{VPD}}{2 (g_1 + \sqrt{VPD})^2}\right) \right)\]

    \begin{itemize}
    \item $c_p$ and $R*$ are constants
    \item $R_{air}$, $\gamma$, and $c_s$ are approximately constant (relative to $\sqrt{VPD}$)
    \item uWUE and $g1$ are constants within plant type
    \end{itemize}
    So \textbf{within each plant type}, whether the atmospheric demand (ET increasing with VPD) or plant response (ET decreasing with VPD) dominates is approximately \textbf{just a function of VPD}!
  \end{overprint}    
\end{frame}

@@

\begin{frame}{Summary - When does VPD drive or reduce ET?}
  \begin{itemize}
  \item Theory predicts that each plant type has a \textbf{critical VPD} \textbf{below which ET will decrease} (plant response dominates), and \textbf{above which ET will increase} (atmospheric demand dominates).
  \item For forest sites, environmental VPD approximately straddles the critical VPD.
  \item For shrubs environmental VPD never exceeds the critical VPD.
  \item Theory tested poorly with FLUXNET data for for crops and grass. 
  \item All plant types exhibited a response far below that of PET.
  \item The new uWUE-version of Penman-Monteith we derived could be used as a complement for PET in drought indices over vegetated surfaces.
  \end{itemize}
\end{frame}
